White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2017-09-19 papers/wdmerger_II/paper.tex
Clean up and simplify discussion of other parameters

We implement the white dwarf collision problem in \castro\ in the \wdmerger\ problem
directory. We prepare the domain in the way most previous papers have done so:
the white dwarf centers of mass are initially separated by a distance of four times
the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
consequently the initial distance, depends on the equation of state used, and this
can vary somewhat depending on the relevant physics used, such as Coulomb corrections
in the Helmholtz equation of state. However, the results do not really depend on the
exact initial distance as long as there is enough time for the WDs to distort in
response to tidal forces as they approach. Their initial velocity is that of
two point-masses in free-fall towards each other coming in from infinity, such that
the contact point is at the origin, and they approach each other along a coordinate axis.
In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
composed of equal parts carbon and oxygen by mass, for comparison to earlier studies;
however, the code does have the ability to vary the impact parameter of the collision
and the composition of the WDs.

@@

The simulation is terminated when the total energy is positive, indicating that
the system has become unbound due to nuclear energy release, and has been
constant or decreasing for the last five timesteps, indicating that no further
meaningful nuclear energy generation is occurring and material is now beginning
to stream off the simulation domain. As all simulations in this paper are of
WD collisions that generate enough energy to unbind the system, this stopping
criterion is applied throughout all simulations shown here.

As in Paper I, the region outside the WDs is filled with a low-density ambient
material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition is the
same as that of the WDs. Reactions in this region are unimportant, so for
computational efficiency, we disable all nuclear burning for zones that have
$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. The initial ambient
temperature is $10^7\ \text{K}$, which is the same as the initial temperature
everywhere throughout the WDs, and we set the temperature floor to the same
value (any zone that reaches below this temperature will be reset).

@@

\subsection{Other Parameters}

We tested a few other parameters that ended up having no serious effect on the outcome.
The dual energy hydrodynamic formalism (discussed in Paper I) has parameters that determine
whether we obtain the internal energy from the total energy or from a separately advected
internal energy variable. This is usually employed when calculating the pressure during
a hydrodynamics update, and we have also added to \castro\ a similar switch when calculating
the initial temperature for a burn. We find that the value of the dual energy parameter has
only a minor effect on the nickel production, yielding variations on the order of a few percent
with no systematic trend. We also tested whether our choice for the temperature floor mattered
by lowering it to $10^6\ \text{K}$ and $10^5\ \text{K}$; this had a minor effect on nickel
production, again on the order of a few percent. It does allow parts of the WDs to
become colder as they move, due to numerical errors in the advecting flow (the Helmholtz
equation of state is very insensitive to temperature, so even small changes in the internal
energy can create very large changes in the implied temperature), but the amount of energy
injected by the collision is so large that the final result is largely insensitive to whatever
the initial temperature of the
WDs were. We also checked whether there was any consequence of our choice to disable the
burning for small enough $\rho$ and $T$. Lowering the minimum $\rho$ for reactions to
$10^0\ \text{g cm}^{-3}$ from $10^6\ \text{g cm}^{-3}$ had a negligible effect,
and varying the minimum $T$ for reactions from $2 \times 10^{7}\ \text{K}$ to
$4 \times 10^{8}\ \text{K}$ also had no meaningful effect. Finally, we also checked
that the 2D axisymmetric simulation matched the results of an equivalent 3D Cartesian simulation.

%TODO: maybe show an image of the temperature field in the WDs before colliding to demonstrate the point about numerical error?

>>>>

We implement the white dwarf collision problem in \castro\ in the \wdmerger\ problem
directory. We prepare the domain in the way most previous papers have done so:
the white dwarf centers of mass are initially separated by a distance of four times
the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
consequently the initial distance, depends on the equation of state used, and this
can vary somewhat depending on the relevant physics used, such as Coulomb corrections
in the Helmholtz equation of state. However, the results do not really depend on the
exact initial distance as long as there is enough time for the WDs to distort in
response to tidal forces as they approach. Their initial velocity is that of
two point-masses in free-fall towards each other coming in from infinity, such that
the contact point is at the origin, and they approach each other along a coordinate axis.
In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
composed of equal parts carbon and oxygen by mass, for comparison to earlier studies;
however, the code does have the ability to vary the impact parameter of the collision
and the mass and composition of the WDs.

@@

We terminate the simulation at $t = 10$ seconds, as we have found that for this
particular setup, all of the significant burning has completed by this time. At
this point, the total energy on the domain has become positive, typically about
$10^{51}$ ergs, and has stopped increasing. This means that the system has become
unbound due to nuclear energy release, and that no further meaningful nuclear energy
generation is occurring.

As in Paper I, the region outside the WDs is filled with a low-density ambient
material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition is the
same as that of the WDs. Reactions in this region are unimportant, so for
computational efficiency, we disable all nuclear burning for zones that have
$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
the results do not depend on this choice. The initial ambient
temperature is $10^7\ \text{K}$, which is the same as the initial temperature
everywhere throughout the WDs, and we set the temperature floor to the same
value (any zone that reaches below this temperature will be reset). Setting the
temperature floor to a lower value has only minor effects on the simulation outcome.

@@
